Molecular dynamics simulation of synchronization in driven particles

### 2021-02-22 manuscript/manuscript_ajp_2020.tex
discovered frequency regimes

%FROM JUNIPER2015
%l¼3.5 mm and a laser power per trap of WE0.75mW (trap stiffness: k=3.8\times10^7 kg /s^2; trap depth: V0 = 75 k_B T
%ref51
%Juniper, M. P. N., Besseling, R., Aarts, D. G. A. L. & Dullens, R. P. A. Acousto- optically generated potential energy landscapes: potential mapping using colloids under flow. Opt. Express 27, 28707–28716 (2012).
%U(x) measured in ratio of k_BT from -75 to -25
%0 < vAC < 8 micrometer/s and frequency f=0.25 to 0.75 Hz.
%0 < vDC < 5 micrometer/s
%convert to force units?
%check when steps exist - in Juniper - steps don't exist for n=1or n=3, even though steps exist for n¼0,2,4,5 and 6.
%flexible chains
%dipole interaction $\Gamma \propto B^2$
%\Gamma =15 (B=0.43mT) and a stiff chain at \Gamma=392 (B=2.2mT).
%Kink
%16 particles in a strong potential energy landscape of 15 minima (trap spacing: l¼5.5 mm; laser power per trap: WE1.75 mW; trap stiffness: k¼8.7?10?7 kg s?2; trap depth: V0¼200 kBT),

%Volpe AJP Parameters
%We will consider a silica microparticle in water with radius R=1micro m, mass m=11picograms, viscosity eta=0.001 Ns/m^2; $\gamma=6\pi \eta R$,  T=300K, and \tau=0.6\micro s.
%thermofisher off the shelf paramagnetic beads
%https://www.thermofisher.com/us/en/home/references/protocols/proteins-expression-isolation-and-analysis/protein-isolation-protocol/m-270-carboxylic-acid.html
%~2\times10^9 beads per 30 mg -> 1.5 \times 10^-11 g, 15 ng

@@

{\it NOTE: YOUR $A_P$ VALUES ARE TOO SMALL TO MAKE THIS A MEANINGFUL PHASE VARIALBE.  YOU NEED TO SCALE WITH THE DRIVE FORCE AMPLITUDES.  THIS CAN ALSO HELP YOU UNDERSTAND WHY YOU ARE GETTING BIG JUMPS IN POSITION COMPARED TO JUNIPER ET AL.  YOU ARE WORKING IN THE WEAK SUBSTRATE LIMIT.  also insert plots of $y/\lambda$ over a single period with the phase plots - like juniper}
  It is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot
  to distinguish fully and partially synchronized
  behavior.  
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  In a system driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\langle v_y \rangle t)/\lambda
  \end{equation}
  centered about the average particle velocity $\langle v_y \rangle$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  In phase diagrams we plot $\phi(t)$ %phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\langle v_y \rangle /\lambda.
  \end{equation}
  %Juniper %REF 53
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  phase space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a 1:1 mode appears as a circle or oval.
  Nodes appear in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is not fully phase locked
  will apppear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized.
  %and the shifts in phase space represent
  %small changes in system behavior.
  With a long simulation time,
  a quasiperiodic phase diagram
  fills the plotting region due to the gradual shift
  in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

  %OKAY, WHY AREN'T MY AMPLITIDES \phi \pm 2\pi and \dot{\phi} \pm 4\pi.
  In Fig.~\ref{fig:phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing substrate amplitude $A_p$.
  We fix the driving force with parameters
  $F^{dc} = 0.1$ and frequency $f = 0.01$ and $F^{ac} = 0.2$.
  In Fig.~\ref{fig:phase}(a)
  there is no substrate ($A_p = 0.0$)
  and the phase plot is a closed oval
  centered at
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$ %ANALYTIC
  and semiminor axis of $\Delta \dot{\phi} = 2\pi v_{max} /\lambda$.
  [insert analysis.]
  As the force from the substrate
  increases,
  nodes form in the phase diagram
  and the motion becomes quasiperiodic
  so the loops in phase space do not fully overlap.
  In Fig.~\ref{fig:phase}(b)
  $A_p = 0.05$
  the spatial properties are similar to
  those in $A_p = 0.0$
  with 
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$.
  The dynamical properties
  are asymmetric,
  with three nodes in the positive $\dot{\phi}$ values.
  The thickening of the line
  indicates the system is not fully synchronized.
  [insert analysis.]
  As the substrate
  force becomes larger,
  $\phi$ shifts to lower values
  and $\dot{\phi}$ to positive values,
  indicating
  the particle can only move 
  backward over a short distance.
  In Fig.~\ref{fig:phase}(c)
  $A_p = 0.1$
  the phase properties are similar to
  those in $A_p = 0.05$,
  the three nodes have 
  increased asymmetry in both phase variables 
  $(\phi,\dot{\phi})$.
  The nodes deepen,
  indicating [insert analysis.].
  In Fig.~\ref{fig:phase}(d)
  $A_p = 0.2$
  the pattern shifts
  to two nodes,
  [indicating change analysis].
  In Fig.~\ref{fig:phase}(e)
  $A_p = 0.25$
  the system returns to a nodeless
  orbit,
  shaped like a triangle [because why?]
  In Fig.~\ref{fig:phase}(f)
  $A_p = 0.3$
  the particle is trapped between substrate minima,
  and traces out a small unclosed
  orbit in phase space [because...]
  The phase velocity is near zero
  throughout the motion -
  [comment on initial conditions]
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{phase.pdf}
      \caption{
        Phase plot of a particle with the same values of $F^{dc}$ and frequency as Fig.~\ref{fig:0}.  $F^{ac} = 0.2$ and $A_p = $ (a) $0.0$, (b) $0.05$, (c) $0.1$, (d) $0.2$, (e) $0.25$ and (f) $0.3$.
      }
      \label{fig:phase}
    \end{figure}
  \end{center}

@@

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{D}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics \cite{Bak1986}.
We note the individual steps
are not strictly horizontal 
since particles move continuously
through the landscape, so changes in $F^{dc}$
will create linear changes in 
net particle displacement that are observable
within a single step.
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{fig3.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$
  as a function of $F^{dc}$, 
  the constant parameter of $F^D(t)$ defined in Eq.~\ref{eq:drive}.
  We fix
  $F^{ac}=0.0$ (blue) and 
  $F^{ac}=0.05$ (orange) with $f = 0.01$ 
  as in Fig.~\ref{fig:0}.}
\label{fig:1}
\end{figure}
\end{center}
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.
%This synchronization
%of the particle velocity can be measured
%in terms of the landscape period.
%Phase locked steps occur for average displacements integer
%multiples of the substrate period $n\lambda$.
%\subsection{Phase plots}
%  \label{ex:phase}

>>>>

%FROM JUNIPER2015
%l¼3.5 mm and a laser power per trap of WE0.75mW (trap stiffness: k=3.8\times10^7 kg /s^2; trap depth: V0 = 75 k_B T
%ref51
%Juniper, M. P. N., Besseling, R., Aarts, D. G. A. L. & Dullens, R. P. A. Acousto- optically generated potential energy landscapes: potential mapping using colloids under flow. Opt. Express 27, 28707–28716 (2012).
%U(x) measured in ratio of k_BT from -75 to -25
%0 < vAC < 8 micrometer/s and frequency f=0.25 to 0.75 Hz.
%0 < vDC < 5 micrometer/s
%convert to force units?
%check when steps exist - in Juniper - steps don't exist for n=1or n=3, even though steps exist for n¼0,2,4,5 and 6.
%flexible chains
%dipole interaction $\Gamma \propto B^2$
%\Gamma =15 (B=0.43mT) and a stiff chain at \Gamma=392 (B=2.2mT).
%Kink
%16 particles in a strong potential energy landscape of 15 minima (trap spacing: l¼5.5 mm; laser power per trap: WE1.75 mW; trap stiffness: k¼8.7?10?7 kg s?2; trap depth: V0¼200 kBT),
%JUNIPER2017
%The driving force is provided by a PI-542.2CD piezo-stage, controlled using the LabView interface, at driving 0.4 ≤ FAC/ζ ≤ 14 µm s−1, and a frequency of 1 velocities of 0.05 ≤ FDC/ζ ≤ 8 µm s−1. AC driving veloc
%Volpe AJP Parameters
%We will consider a silica microparticle in water with radius R=1micro m, mass m=11picograms, viscosity eta=0.001 Ns/m^2; $\gamma=6\pi \eta R$,  T=300K, and \tau=0.6\micro s.
%thermofisher off the shelf paramagnetic beads
%https://www.thermofisher.com/us/en/home/references/protocols/proteins-expression-isolation-and-analysis/protein-isolation-protocol/m-270-carboxylic-acid.html
%~2\times10^9 beads per 30 mg -> 1.5 \times 10^-11 g, 15 ng

@@

%{\it NOTE: YOUR $A_P$ VALUES ARE TOO SMALL TO MAKE THIS A MEANINGFUL PHASE VARIALBE.  YOU NEED TO SCALE WITH THE DRIVE FORCE AMPLITUDES.  THIS CAN ALSO HELP YOU UNDERSTAND WHY YOU ARE GETTING BIG JUMPS IN POSITION COMPARED TO JUNIPER ET AL.  YOU ARE WORKING IN THE WEAK SUBSTRATE LIMIT.  also insert plots of $y/\lambda$ over a single period with the phase plots - like juniper}
  It is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot
  to distinguish fully and partially synchronized
  behavior.  
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  In a system driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\langle v_y \rangle t)/\lambda
  \end{equation}
  centered about the average particle velocity $\langle v_y \rangle$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  In phase diagrams we plot $\phi(t)$ %phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\langle v_y \rangle /\lambda.
  \end{equation}
  %Juniper %REF 53
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  phase space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a 1:1 mode appears as a circle or oval.
  Nodes appear in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is not fully phase locked
  will apppear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized.
  %and the shifts in phase space represent
  %small changes in system behavior.
  With a long simulation time,
  a quasiperiodic phase diagram
  fills the plotting region due to the gradual shift
  in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

  %OKAY, WHY AREN'T MY AMPLITIDES \phi \pm 2\pi and \dot{\phi} \pm 4\pi.
  In Fig.~\ref{fig:phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing substrate amplitude $A_p$.
  We fix the driving force with parameters
  $F^{dc} = 0.1$ and frequency $f = 0.01$ and $F^{ac} = 0.2$.
  In Fig.~\ref{fig:phase}(a)
  there is no substrate ($A_p = 0.0$)
  and the phase plot is a closed oval
  centered at
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$ %ANALYTIC
  and semiminor axis of $\Delta \dot{\phi} = 2\pi v_{max} /\lambda$.
  [insert analysis.]
  As the force from the substrate
  increases,
  nodes form in the phase diagram
  and the motion becomes quasiperiodic
  so the loops in phase space do not fully overlap.
  In Fig.~\ref{fig:phase}(b)
  $A_p = 0.05$
  the spatial properties are similar to
  those in $A_p = 0.0$
  with 
  $(\phi,\dot{\phi})=(10,0)$
  with a semimajor axis of $\Delta \phi = 20$.
  The dynamical properties
  are asymmetric,
  with three nodes in the positive $\dot{\phi}$ values.
  The thickening of the line
  indicates the system is not fully synchronized.
  [insert analysis.]
  As the substrate
  force becomes larger,
  $\phi$ shifts to lower values
  and $\dot{\phi}$ to positive values,
  indicating
  the particle can only move 
  backward over a short distance.
  In Fig.~\ref{fig:phase}(c)
  $A_p = 0.1$
  the phase properties are similar to
  those in $A_p = 0.05$,
  the three nodes have 
  increased asymmetry in both phase variables 
  $(\phi,\dot{\phi})$.
  The nodes deepen,
  indicating [insert analysis.].
  In Fig.~\ref{fig:phase}(d)
  $A_p = 0.2$
  the pattern shifts
  to two nodes,
  [indicating change analysis].
  In Fig.~\ref{fig:phase}(e)
  $A_p = 0.25$
  the system returns to a nodeless
  orbit,
  shaped like a triangle [because why?]
  In Fig.~\ref{fig:phase}(f)
  $A_p = 0.3$
  the particle is trapped between substrate minima,
  and traces out a small unclosed
  orbit in phase space [because...]
  The phase velocity is near zero
  throughout the motion -
  [comment on initial conditions]
  \begin{center}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{phase.pdf}
      \caption{
        Phase plot of a particle with the same values of $F^{dc}$ and frequency as Fig.~\ref{fig:0}.  $F^{ac} = 0.2$ and $A_p = $ (a) $0.0$, (b) $0.05$, (c) $0.1$, (d) $0.2$, (e) $0.25$ and (f) $0.3$.
        [first 100 timesteps removed to reduce transients]
      }
      \label{fig:phase}
    \end{figure}
  \end{center}

@@

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{D}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics \cite{Bak1986}.
We note the individual steps
are not strictly horizontal 
since particles move continuously
through the landscape, so changes in $F^{dc}$
will create linear changes in 
net particle displacement that are observable
within a single step.
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{fig3.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$
  as a function of $F^{dc}$, 
  the constant parameter of $F^D(t)$ defined in Eq.~\ref{eq:drive}.
  We fix
  $F^{ac}=0.0$ (blue) and 
  $F^{ac}=0.05$ (orange) with $f = 0.01$ 
  as in Fig.~\ref{fig:0}.
  {TODO: with these parameters, the steps are fractional!}
}
\label{fig:1}
\end{figure}
\end{center}
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.
%This synchronization
%of the particle velocity can be measured
%in terms of the landscape period.
%Phase locked steps occur for average displacements integer
%multiples of the substrate period $n\lambda$.
%\subsection{Phase plots}
%  \label{ex:phase}
